---
title: "Challenge1NYU"
author: "Bruce Kusko"
date: "January 30, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, echo=FALSE, message=FALSE}
library(dplyr)
```

## Challenge 1 
### NYU Data Science Workshop

```{r}
## read in the data set
hony <- read.csv("./humansofnewyork.csv", stringsAsFactors=FALSE)
```


Do some data exploration:
```{r}
dim(hony)
head(hony,3)
summary(hony)
```


#### 1. How many status updates have been posted on this page?

If a status update can be either a like, a comment, or a share, then each row is a status update

```{r, echo=FALSE}
## length of data frame is first dimension
ans1 <- dim(hony)[1]
print(paste0(ans1, ' status updates have been posted on this page', sep=''))
```

#### 2. What is the total number of likes, comments, and shares it received?

Here we just need to sum the three columns and then get grand total.

```{r}
ans2a <- colSums(hony[,8:10])
print(paste0('Total likes = ', ans2a[1], 'total comments = ', ans2a[2], 'and total shares = ', ans2a[3]))
print(paste0(' So the total of totals is: ', ans2a[1] + ans2a[2] + ans2a[3]))
```


#### 3. What is the content of the post with the highest number of shares?

Looks like they're asking for the "message" from the row with the most "shares_count".

```{r, echo=FALSE}
## message is the 3rd column
ans3 <- head(arrange(hony, desc(shares_count)),1)[,3]
print(paste0('The "message" with the most shares is: ', ans3))
```


#### 4. What was the date in which the first photo was posted?

We see from hony$type that there are links, photos, videos, and statuses.
We want to find the first 'update' with a photo
?? Subset the rows with photos and take the earliest one??
Problem is, 'created time' is not in standard format, it has a T in the middle, and is a char.
Create a true 'date' column (called 'created_time2')

```{r}
# convert the 'created_time' col to a date
# first strip out the YYYY-MM-DD into a new column
hony$created_time2 <- substr(hony$created_time, 1, 10)
## now change that 'char' to a 'date'
hony$created_time2 <- as.Date(hony$created_time2)
```
Sort by oldest
```{r}
ans4 <- head(arrange(hony, created_time2),1)[,11]
print(paste0('The date of the oldest posting with a photo is: ', ans4))
```

#### 5. What is the total number of likes that the page has ever received, excluding its most popular post?

Not sure, but it sounds like we want the total likes minus the most popular (by likes) post.
So need to sort, descending, and take the sum of the 2nd thru end.




end as of now

